---
title: "R Notebook"
output: html_notebook
---
#loading packages
```{r, include = F}
library(rstatix) #pipe-friendly statistics functions
library(ggpubr) #easily publication ready plots
library(plotrix)#for standard error of the mean function
library(tidyverse) #includes dplyr, ggplot, readr, tidyr, etc
library(ggsignif) #can plot significance values on graphs
library(graphics)
```

#Loading metadata files
```{r}
colorscore_metadata <- read.csv("../colorscore_metadata.csv") #Supplemental Table 2, "metadata" sheet
pam <- read.csv("../pam.csv") #Supplemental Table 2, "PAM" sheet
Rintensity <- read.csv("../Rintensity_data.csv") #Supplemental Table 2, "R-intensity" sheet
```

# Experiment Timeline

**Coral collections and fragmentation**: February 6-8, 2019
**Field control heat-stress assay**: March 15, 2019 - April 10, 2019
**Stress-hardening temperature treatment (laboratory control vs. variable)**: March 21, 2019 - June 17, 2019
**2nd Heat-stress assay**: June 17, 2019 - July 9, 2019

# Sample Sizes
```{r}
colorscore_metadata %>% 
  select(Puck.ID, Treatment, Genotype) %>% 
  distinct() %>% 
  group_by(Treatment, Genotype) %>% 
  dplyr::mutate(samplesize = n()) %>% 
  distinct(samplesize) %>%
  arrange(Treatment)

#sample size of field control
colorscore_metadata %>% 
  select(Puck.ID, Treatment, Genotype) %>% 
  distinct() %>% 
  filter(Treatment =="Field control") %>% 
  group_by(Treatment) %>% 
  dplyr::mutate(samplesize = n()) #46 fragments 

#number used in stress hardening
colorscore_metadata %>% 
  select(Puck.ID, Treatment, Genotype, reasonPulled) %>% 
  distinct() %>% 
  filter(!Treatment =="Field control") %>% 
  group_by(Treatment) %>% 
  dplyr::mutate(samplesize = n()) #82 fragments
#39 variable, 43 laboratory control
```

# Heat-stress assay Figures
```{r}
heatstressassays <- read_csv("../heatstressassays.csv") #Supplemental Table 1, "Raw Heat-Stress Assay Log" sheet
heatstressassays$Date<- lubridate::mdy(heatstressassays$Date)

heatstressassays %>% 
  filter(Date <= "2019-04-10") %>% 
  filter(Date >= "2019-03-15") %>% 
  group_by(Date, Time) %>% 
  summarise(mean = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(data = .,
       aes(x = Date, y = mean)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.2) +
  theme_classic() +
  labs(title = "Field control heat-stress assay",
       y = "Temperature (°C)",
       x = "Date")

heatstressassays %>% 
  filter(!Experiment =="Field") %>% 
  group_by(Date, Time) %>% 
  summarise(mean = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(data = .,
       aes(x = Date, y = mean)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.2) +
  theme_classic() +
  labs(title = "Laboratory control/variable heat-stress assay",
       y = "Temperature (°C)",
       x = "Date")
```
#temperature treatment figure
```{r}
temptreatment <- read_csv("../temptreatment.csv")

temptreatment %>% 
group_by(Time, Treatment) %>% 
  summarise(avg_temp = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp, color = Treatment)) + 
  theme_classic() +
  geom_ribbon(aes(fill=Treatment, group = Treatment, ymin = avg_temp - sd, ymax = avg_temp + sd), alpha = 0.2) +
   geom_line(aes(group = Treatment)) +
  labs(y="Temperature (C)")
```

#health of corals going into heat-stress assay
```{r}
colorscore_metadata$colorscoredate <- lubridate::mdy(colorscore_metadata$colorscoredate)
colorscore_metadata  %>% 
  filter(colorscoredate == "2019-03-18" | colorscoredate == "2019-06-17") %>% 
  group_by(Treatment, Genotype) %>% 
  summarise(avg_score = mean(Score), sd_score = sd(Score), n = n())

colorscore_metadata  %>% 
  filter(colorscoredate == "2019-03-18" | colorscoredate == "2019-06-17") %>% 
  group_by(Treatment) %>% 
  summarise(avg_score = mean(Score), sd_score = sd(Score), n = n())
```
#difference in number of days among treatment groups 
```{r}
colorscore_metadata %>% 
  group_by(Treatment) %>%
  summarise(averagedays = mean(daysUntilPulled), sd = sd(daysUntilPulled)) %>%
  ggplot(., aes(x = Treatment, y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - sd, ymax = averagedays + sd), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  scale_fill_manual(values = c("#60DBDB","#00E281",  "#F54A34"))
```

#each genotype number of days until bleaching or RTL
```{r}
colorscore_metadata %>% 
  group_by(Genotype, Treatment) %>%
  summarise(averagedays = mean(daysUntilPulled), sd = sd(daysUntilPulled)) %>%
  ggplot(., aes(x = Genotype, y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - sd, ymax = averagedays + sd), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  scale_fill_manual(values = c("#60DBDB","#00E281",  "#F54A34"))
ggsave("genotypes_sd.pdf")

colorscore_metadata %>% 
  select(Puck.ID, Treatment, Genotype, reasonPulled) %>% 
  distinct() %>% 
  group_by(Genotype, Treatment, reasonPulled) %>% 
  dplyr::mutate(count = n()) %>%
  distinct(count) %>% 
  ggplot(., aes(x=Treatment, y=count, fill = reasonPulled)) +
  geom_bar(stat="identity", color = "black") + 
  facet_wrap(~ Genotype) + theme_classic() + 
  labs(fill = "Reason Pulled", x = "Treatment", y = "Number of Corals") +
  scale_fill_manual(labels=c("Bleached", "Tissue Loss"), values = c("white", "brown")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Two-Way ANOVA for genotype and treatment
```{r}
colorscore_metadata %>% 
  select(Puck.ID:daysUntilPulled) %>% 
  distinct() %>% 
  aov(daysUntilPulled ~ Genotype + Treatment + Genotype*Treatment, data = .) -> anova_two_way

plot(anova_two_way)

summary(anova_two_way)
TukeyHSD(anova_two_way) -> tukey_anova_genotreatment

as.data.frame(tukey_anova_genotreatment$OriginalTreatment) %>% 
  write.csv("tukey_anova_genotreatment.csv",  row.names = T)

plot(TukeyHSD(anova_two_way))

colorscore_metadata %>% 
  select(Puck.ID, Treatment, daysUntilPulled, Genotype) %>% 
  distinct() %>% 
    rstatix::levene_test(daysUntilPulled ~ Genotype) #not significant so variance is good 
  
shapiro_test(anova_two_way$residuals) #not significant, so data is normal
```


#grouping by reef site (North, Mid, South)
```{r}
colorscore_metadata %>% 
  mutate(ecoregion = case_when(Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  dplyr::select(Puck.ID, Treatment, daysUntilPulled, ecoregion, Genotype) %>% 
  group_by(ecoregion, Treatment) %>% 
dplyr::summarise(averagedays = mean(daysUntilPulled), sd = sd(daysUntilPulled)) %>%
  ggplot(., aes(x = factor(ecoregion, levels = c("North", "Mid", "South")), y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - sd, ymax = averagedays + sd), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  scale_fill_manual(values = c("#60DBDB","#00E281",  "#F54A34"))

colorscore_metadata %>% 
  select(Puck.ID, Treatment, Genotype, reasonPulled) %>% 
  distinct() %>% 
  mutate(ecoregion = case_when(Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  dplyr::group_by(ecoregion, Treatment, reasonPulled) %>% 
  dplyr::mutate(count = n()) %>% 
  distinct(count) %>% 
  ggplot(., aes(x=ecoregion, y=count, fill = reasonPulled)) +
  geom_bar(stat="identity", color = "black") + 
  facet_wrap(~ Treatment) + theme_classic() + 
  labs(fill = "Reason Pulled", x = "Treatment", y = "Number of Corals") +
  scale_fill_manual(labels=c("Bleached", "Tissue Loss"), values = c("white", "brown")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Two-Way ANOVA for original reef and treatment
```{r}
colorscore_metadata %>% 
  mutate(ecoregion = case_when(Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  dplyr::select(Puck.ID, Treatment, daysUntilPulled, ecoregion) %>% 
  distinct() %>% 
  aov(daysUntilPulled ~ ecoregion + Treatment + ecoregion*Treatment, data = .) -> region_anova
  
summary(region_anova)
  
colorscore_metadata %>% 
  mutate(ecoregion = case_when(Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  dplyr::select(Puck.ID, Treatment, daysUntilPulled, ecoregion) %>%
  distinct() %>% 
    rstatix::levene_test(daysUntilPulled ~ ecoregion) #not significant so variance is good 

shapiro_test(region_anova$residuals) #not significant, so data is normal
```

# Chi-squared contingency tables
```{r}
#comparison of treatment groups
colorscore_metadata  %>% 
  dplyr::select(Puck.ID:reasonPulled) %>% 
  distinct() %>% 
  dplyr::group_by(Treatment, reasonPulled) %>% 
  dplyr::summarise(count = n()) %>% 
  pivot_wider(names_from = reasonPulled, values_from = count) %>% 
  column_to_rownames(., var = "Treatment") -> contingencytable

chisq_test(contingencytable, correct = F) #correct = F gets rid of Yate's contingency correction

colorscore_metadata %>% 
  select(Puck.ID, Treatment, reasonPulled) %>% 
  distinct() %>% 
  mosaicplot(~ Treatment + reasonPulled, data = ., color = 8:9)
```

```{r}
#all treatment groups condensed
colorscore_metadata  %>% 
  mutate(ecoregion = case_when(Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  select(Puck.ID, Treatment, reasonPulled, ecoregion) %>% 
  distinct() %>% 
  group_by(ecoregion, reasonPulled) %>% 
  dplyr::summarise(count = n()) %>% 
  pivot_wider(names_from = reasonPulled, values_from = count) %>% 
  column_to_rownames(., var = "ecoregion") -> contingencytable_alltreatments_ecoregion

chisq_test(contingencytable_alltreatments_ecoregion, correct = F) 

#mosaic plot
colorscore_metadata %>% 
  mutate(ecoregion = case_when(Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  mutate(reorder_levels(., ecoregion, c("North", "Mid", "South"))) %>% 
  select(Puck.ID, Treatment, reasonPulled, ecoregion) %>% 
  distinct() %>% 
  mosaicplot(~ ecoregion + reasonPulled, data = ., color = 8:9)

```

```{r}
#comparison of original reef site (using laboratory control and variable groups)

colorscore_metadata  %>% 
  filter(!Treatment == "Field control") %>% 
  mutate(ecoregion = case_when(Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  select(Puck.ID, Treatment, reasonPulled, ecoregion) %>% 
  distinct() %>% 
  group_by(ecoregion, reasonPulled) %>% 
  dplyr::summarise(count = n()) %>% 
  pivot_wider(names_from = reasonPulled, values_from = count) %>% 
  column_to_rownames(., var = "ecoregion") -> contingencytable_ecoregion

chisq_test(contingencytable_ecoregion, correct = F) 

#mosaic plot
colorscore_metadata %>% 
  filter(!Treatment == "Field control") %>% 
  mutate(ecoregion = case_when(Genotype == "Mid A" | Genotype == "Mid B" ~ "Mid",
                               Genotype == "North A" | Genotype == "North B" ~ "North",
                               Genotype == "South A" | Genotype == "South B" ~ "South")) %>% 
  mutate(reorder_levels(., ecoregion, c("North", "Mid", "South"))) %>% 
  select(Puck.ID, Treatment, reasonPulled, ecoregion) %>% 
  distinct() %>% 
  mosaicplot(~ ecoregion + reasonPulled, data = ., color = 8:9)
```


#PAM Analysis
```{r}

pam$pam_date <- lubridate::ymd(pam$pam_date)
pam %>% 
  filter(!OriginalTreatment=="Field") %>% 
  group_by(OriginalTreatment, pam_date) %>% 
  mutate(mean_fvfm = mean(pam_fvfm),sd_fvfm = sd(pam_fvfm)) %>% 
  ggplot(.,
       aes(
         x = pam_date,
         y = mean_fvfm,
         color = OriginalTreatment
       )) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_fvfm - sd_fvfm, ymax = mean_fvfm + sd_fvfm)) +
  labs(x = "Date",
       y = "fv/fm",
       color = "Original Treatment") +
  theme(text = element_text(size = 13)) +
 scale_color_manual(labels=c("Laboratory control", "Variable"), values = c( "#60DBDB", "#F54A34"))+  #"Field" = "#00E281"
 scale_x_date(date_labels = "%d %b", date_breaks = "2 week") +
  theme_classic()


pam %>% 
  filter(OriginalTreatment=="Field") %>% 
  group_by(OriginalTreatment, pam_date) %>% 
  mutate(mean_fvfm = mean(pam_fvfm),sd_fvfm = sd(pam_fvfm)) %>% 
  ggplot(.,
       aes(
         x = pam_date,
         y = mean_fvfm,
         color = OriginalTreatment
       )) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_fvfm - sd_fvfm, ymax = mean_fvfm + sd_fvfm)) +
  labs(x = "Date",
       y = "fv/fm",
       color = "Original Treatment") +
  theme(text = element_text(size = 13)) +
 scale_color_manual(labels=c("Field control"), values = c( "#00E281"))+  #"Field" = "#00E281"
 scale_x_date(date_labels = "%d %b", date_breaks = "2 week") +
  theme_classic()
```

R intensity
```{r}
Rintensity$Date <- lubridate::mdy(Rintensity$Date)

Rintensity %>%
  filter(Date <= "2019-07-10") %>% 
  group_by(Date, Treatment) %>%
  dplyr::summarise(average = mean(Average.R.Intensity), sd = sd(Average.R.Intensity)) %>% 
  ggplot(., aes(x=Date, y=average, color=Treatment)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = average - sd, ymax = average + sd), width = 0.5) +
  theme_classic() +
  theme(text = element_text(size = 13)) +
  scale_color_manual(labels=c("Laboratory control", "Variable"), values = c("#60DBDB", "#F54A34")) +
  labs(y="R Intensity") +
  scale_y_reverse()
ggsave("rintensity.pdf")
```

